---
title: "Homework 6"
author: "Rachel Hall"
date: "Feb 21, 2024"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2) #for graphics 
library(MASS)
library(dplyr) 
```

<!-- #### Used Li_Soil_Dataset from examples -->
<!-- z <- rnorm(n=3000,mean=0.2) -->
<!-- z <- data.frame(1:3000,z) -->
<!-- names(z) <- list("ID","myVar") -->
<!-- z <- z[z$myVar>0,] -->
<!-- str(z) -->
<!-- summary(z$myVar) -->

#### Reading Data:
```{r echo=TRUE,results='hide'}
z<-read.csv("Li_Soil_Dataset.csv", sep=",", comment.char = '#')
str(z)
z<-na.omit(z)
z$myVar<-z$Plant.biomass..g.
summary(z)
```

#### Plotting Histogram:
```{r message=FALSE}
p1 <- ggplot(data=z, aes(x=myVar, y=after_stat(density))) +
  geom_histogram(color="grey",fill="cadetblue3",linewidth=0.2) 
```

##### Adding Empirical Density Curve:
```{r echo=TRUE, message=FALSE}
p1 <-  p1 +  geom_density(linetype="dotted",linewidth=0.75)
print(p1)
```

##### Getting maximum likelihood parameters for normal:
```{r echo=TRUE, results='hide'}
normPars <- fitdistr(z$myVar,"normal")
print(normPars)
str(normPars)
```
```{r echo=TRUE}
normPars$estimate["mean"] # note structure of getting a named attribute
```

#### Plot Normal Probability:
``````{r echo=TRUE, message=FALSE}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]

xval <- seq(0,max(z$myVar),len=length(z$myVar))

stat <- stat_function(aes(x = xval, y = after_stat(y)), fun = dnorm, colour="mediumorchid", n = length(z$myVar), args = list(mean = meanML, sd = sdML))
p1 + stat
```

#### Plot exponential probability density:
```{r echo=TRUE, message=FALSE}
expoPars <- fitdistr(z$myVar,"exponential")
rateML <- expoPars$estimate["rate"]

stat2 <- stat_function(aes(x = xval, y = after_stat(y)), fun = dexp, colour="green", n = length(z$myVar), args = list(rate=rateML))
p1 + stat + stat2
```

#### Plot uniform probability density:
```{r echo=TRUE, message=FALSE}
stat3 <- stat_function(aes(x = xval, y = after_stat(y)), fun = dunif, colour="darkgreen", n = length(z$myVar), args = list(min=min(z$myVar), max=max(z$myVar)))
p1 + stat + stat2 + stat3
```

#### Plot gamma probability:
```{r echo=TRUE, message=FALSE, warning=FALSE}
gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = after_stat(y)), fun = dgamma, colour="navy", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p1 + stat + stat2 + stat3 + stat4
```

#### Plot beta probability:
```{r echo=TRUE, message=FALSE, warning=FALSE}
pSpecial <- ggplot(data=z, aes(x=myVar/(max(myVar + 0.1)), y=after_stat(density))) +
  geom_histogram(color="grey60",fill="plum1",linewidth=0.2) + 
  xlim(c(0,1)) +
  geom_density(linewidth=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$myVar/max(z$myVar + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = after_stat(y)), fun = dbeta, colour="orchid4", n = length(z$myVar), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

#### Comparing Likelihood (higher number = best fit):
```{r echo=TRUE}
normPars$loglik 
gammaPars$loglik 
expoPars$loglik
```

##### Gamma distribution would be the best fit as it has the highest number for likelihood.

#### Simulating Data:
```{r echo=TRUE,results='hide',message=FALSE, warning=FALSE}
z <- rgamma(n= 60, shape=shapeML, rate=rateML) 
z <- data.frame(1:60,z)

names(z) <- list("ID","myVar")
z <- z[z$myVar>0,]
str(z)
z <- na.omit(z)

p2 <- ggplot(data=z, aes(x=myVar, y=..density..)) +
  geom_histogram(color="snow",fill="darkseagreen",size=0.2) 
print(p2)

normPars <- fitdistr(z$myVar,"normal")
print(normPars)

str(normPars)

gammaPars <- fitdistr(z$myVar,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]
```
```{r echo=TRUE, message=FALSE}
stat5 <- stat_function(aes(x = xval, y = after_stat(y)), fun = dgamma, colour="darkslategrey", n = length(z$myVar), args = list(shape=shapeML, rate=rateML))
p2 + stat5

p1 + stat4
```

#### Comparing the simulated plot to the data we imported:

##### While the simulated data does not exactly match the original data set, it does show similar trends in the values. Both the simulated data and the original data are right skewed as well, which illustrates this similarity. 
