---
title: "Homework 10"
author: "Rachel Hall"
date: "April 3, 2024"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
```

#### Question 1 

##### Using a for loop, write a function to calculate the number of zeroes in a numeric vector. Before entering the loop, set up a counter variable counter <- 0. Inside the loop, add 1 to counter each time you have a zero in the vector. Finally, use return(counter) for the output.


```{r}
number_zs <- function(z = round(runif(100), digits=0)) {
  counter <- 0
  for (i in seq_along(z)) {
    if (z[i]==0) counter <- counter+1}
  return(counter)}
number_zs()
```


#### Question 2 

##### Use subsetting instead of a loop to rewrite the function as a single line of code.
```{r}
number_zs2 <- function(z= round(runif(100), digits = 0)) {
  print(counter <-sum(z==0))}
number_zs2()
```


#### Question 3

##### Write a function that takes as input two integers representing the number of rows and columns in a matrix. The output is a matrix of these dimensions in which each element is the product of the row number x the column number.

```{r}
product_matrix <- function(r, c) {
  
  m <- matrix(NA, nrow = r, ncol = c)
  for (i in 1:nrow(m)) {
    for (j in 1:ncol(m)) {
      m[i,j] <- i*j}}
  print(m)}

product_matrix(r=3, c=5)
```


#### Question 4 

##### 4a. Simulate a dataset with 3 groups of data, each group drawn from a distribution with a different mean. The final data frame should have 1 column for group and 1 column for the response variable.

```{r}

Group <- c(rep("Control",10),rep("Treatment1",10),rep("Treatment2",10))
Data <- c(rnorm(10,mean = 7, sd = 3),
          rnorm(10,mean = 10, sd = 2),
          rnorm(10,mean = 20, sd = 4))
df <- data.frame(Group,Data)

head(df)
```

##### 4b. Write a custom function that 1) reshuffles the response variable, and 2) calculates the mean of each group in the reshuffled data. Store the means in a vector of length 3.


```{r}
shuffle_data <- function(df) {
   df <- data.frame(Group,Data) 
    df[,2]<-sample(df$Data) # reshuffle data
   df<-group_by(df, Group)
  mean<-summarize(df,Mean=mean(Data))
  return(mean)
}
print(MeanTable<-shuffle_data())
print(MeanVals<-MeanTable[,2])
```


##### 4c. Use a for loop to repeat the function in b 100 times. Store the results in a data frame that has 1 column indicating the replicate number and 1 column for each new group mean, for a total of 4 columns.

```{r}
d_new <- data.frame(ID=1:100,
                      Control_Mean=NA,
                      Trt1_Mean=NA,
                      Trt2_Mean=NA)

for (i in 1:100) {
  new_means <- shuffle_data(df)
  d_new[i,2] <- new_means[1,2]
  d_new[i,3] <- new_means[2,2]
  d_new[i,4] <- new_means[3,2]
}
print(d_new)

```

##### 4d. Use qplot() to create a histogram of the means for each reshuffled group. Or, if you want a challenge, use ggplot() to overlay all 3 histograms in the same figure. How do the distributions of reshuffled means compare to the original means?

```{r include=FALSE}
library(ggplot2)
library(ggthemes)
library(patchwork)
```
```{r, message=FALSE}

g1 <- ggplot(data=d_new) + aes(x=Control_Mean) + geom_histogram(fill="lightskyblue",color="cornflowerblue")

g2 <- ggplot(data=d_new) + aes(x=Trt1_Mean) + geom_histogram(fill="darkseagreen1",color="darkseagreen4")

g3 <- ggplot(data=d_new) + aes(x=Trt2_Mean) + geom_histogram(fill="plum1",color="orchid")

```
```{r, message=FALSE}
g1 + g2 + g3 + plot_layout(ncol=1)

```
