---
title: "Homework 4"
author: "Rachel Hall"
date: "2024-02-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
```

## Question 1
```{r echo=TRUE}
x <-1.1
a <-2.2
b <-3.3

```
#### 1a
```{r echo=TRUE}
z <- x^(a^b)
print(z) 

```

#### 1b
```{r echo=TRUE}
z <- (x^a)^b
print(z)
```

#### 1c
```{r echo=TRUE}
z <- (3*(x^3))+(2*(x^2))+1 
print(z)
```

## Question 2
#### 2a
```{r echo=TRUE}
a2 <- c(1:8,seq(7,1,-1))
print(a2)
```

#### 2b
```{r echo=TRUE}
b2 <- seq(1,5)
print(b2)
rep(x=b2,times=b2)
```

#### 2c
```{r echo=TRUE}
b2 <- seq(1,5)
c2 <- seq(5,1)
rep(x=c2,times=b2)
```

## Question 3
```{r echo=TRUE}
set.seed(10)
x2 <- runif(2)
print(x2)
```

#### Finding the hypotenuse
```{r echo=TRUE}
r <- sqrt(x2[1]^2+x2[2]^2)
print(r)

```

#### Finding the angle of the vector (in radians)
```{r echo=TRUE}
theta <- asin(x2[2]/r)
print(theta)

polarcoords <-c(r, theta)
print(polarcoords)
```


## Question 4
```{r echo=TRUE}
queue <- c("sheep", "fox", "owl", "ant")
print(queue)
```

#### 4a: Serpent gets in line 
```{r echo=TRUE}
queue[5] <- ("serpent")
print(queue)
```

#### 4b: The sheep enters the ark (gets out of line)
```{r echo=TRUE}
queue <- queue[-1]
print(queue)
```

#### 4c: The donkey arrives and talks his way to the front of the line.
```{r echo=TRUE}
queue <- c("donkey", queue)
print(queue)
```

#### 4d: The serpent gets impatient and leaves.
```{r echo=TRUE}
queue <- queue[-5]
print(queue)
```

#### 4e: The owl gets bored and leaves.
```{r echo=TRUE}
queue <- queue[-3]
print(queue)
```

#### 4f: The aphid arrives and the ant invites him to cut in line.
```{r echo=TRUE}
queue <- c(queue[1], queue[2], "aphid", queue[3])
print(queue)
```

#### 4g: Finally, determine the position of the aphid in the line.
```{r echo=TRUE}
which(queue == "aphid")
```

## Question 5
```{r echo=TRUE}
vec <- seq(1:100)
print(vec)

vec <- vec[vec %% 2 !=0 & vec%%3 !=0 & vec%%7 !=0]
print(vec)

```





