---
title: "Working with Markdown"
author: "Rachel Hall"
date: "2024-01-23"
output: 
  html_document: 
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
For markdown, use asterisks to *indicate italic text*.



Double line return for a new chunk of text like this. 

Name  
Department of Biology   
University of Vermont

# Hash marks for headers

## more hashes for progressively

### smaller

#### headers

**bold face this way**

>Offset quote which embeds for extended material that you would like to see.

Back to plain text

* first list item
* second item
   - spacing for indents
   
   - ^superscript^ but also a ~subscript~ like this
   
   
 [set up a link with text](website address)
 
 also footnotes^[at bottom of page]
 
First Header  | Second Header
------------- | -------------
Content Cell  | Content Cell
Content Cell  | Content Cell

```
Fencing sets boundaries and everything inside the fencing *is shown as literal text* without rendering
```
$$a + b =c$$
We can also use Latex in line with a single dollar sign $a+b=c$ like this.

### Subscripts in LaTex

$$H_0 = Z_{a + b}$$
$$S=cA^z_1 + z_{2 + x}$$

$$\alpha = \frac{\beta}{\delta + \gamma_x}$$

$$z = \sum_{i=1}^X{K}$$

### “Escaping” the  special character in LaTeX
Dollar sign distinguishes LaTex math expression from regular text. Backslash is used to generate a special symbol or command. 

Use backslash $$\backslash $$
$$\backslash \alpha \le b \backslash$$

### Rendering in plain text
$$P(Occurrence Of Species A) = Z$$

$$P(\mbox{Occurrence Of Species A}) = Z$$

### Embedding 

Easy to embed LaTex in line $\pi + 3$ like so. 
We can also use markdown fencing in line with a single backtick `pi + 3` like this. 
We can even execute r code inside of markdown text `r pi + 3`

### Running R Code 
R chunk is created using three backticks and { r }

Eval is whether or not to evaluate the code and include its results. 

Echo determines if the code if displayed with the results of the code (false = doesn't show).


```{r eval=TRUE, echo=TRUE}
# Use comments extensively in ALL of your coding!
Pred <- seq(1,10)     # make a vector of integers from 1 to 10
Res <- runif(10)      # generate 10 draws from a random uniform (0,1) distribution

# print the random numbers
print(Res)

# plot the graph
plot(x=Pred,y=Res,type="b")
